### 2018-03-07 manuscript/manuscript.Rmd
Morgan's methods section

---
title: "Untitled"
author: "Morgan Hamm"
date: "March 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Nucleotide Diversity Calculation
Nei  defined the nucleotide diversity statistic in his original paper as:
“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
And provided the equation:

$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$

Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.

a more general form, that treats each sequence in the population as unique can be written as:

$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
\qquad
\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
0 \quad \textrm{if} \quad N_{ik} = N_{jk}
\end{array}\right.$

where $N_{ik}$ is the nucelotide (A, T, C or G) at position k on the ith sequence of the population.
$L$ is the lenght of the sequence (the same for all sequences in the population, 
assumes only SNPs exist, no indels). n is the total number of sequences in the population.

from this form we can re-arange summations to the form below:

$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
\qquad
\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$

here $\pi_k$ can be thought of as the site-wise nucleotide diversity at postion k, and is equal to 
the nucleotide diversity of a sequence of lenght 1 at location k. we can calculate $\pi_k$ for each
site, then average those over the sequence lenght to calculate $\pi$, the nucleotide diversity 
of the sequence.

The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant. note, $\pi_k$ is equal to 0 at all locations without variants.
This is also what is displayed in the Diversity Plot tab of the web app.

## Detailed $\pi_k$ calculaiton simplification.

The formula for $\pi_k$ above requires comparing every sequence to everyother sequence at location k,
however, we know there are only a few variant forms at each individual location.

so we can revert back to using Nei's original forula, modifying it slightly,
replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]

$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$

$\pi_{ij(k)} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad i \neq j \\
0 \quad \textrm{if} \quad i = j
\end{array}\right.$

we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.

we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:

$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$

next we substitute our deffinition of $n_{!i}$

$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$

then distribute and split summation:

$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$

finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:

$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$

this simplified form for $\pi_k$ is used by the app, because the counts of unique 
variants at a single nucleotide location can easily be sumaraized in R.