### 2018-07-12 manuscript/07-discussion.Rmd

#Discussion

Analyzing natural diversity within the gene families constituting the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses. This analysis in particular has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes. Despite this relative conservation, this analysis has also identified many accessions containing new alleles of interest in the *AFB* and *Aux/IAA* families. This analysis has also

Importance of 
-AFB
-IAA degron
-TPL
-ARF DBD and PB1