### 2018-07-13 manuscript/07-discussion.Rmd

#Discussion

Our analysis of the natural diversity in the gene families that make up the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses. These results have been generated by a team of undergraduate and graduate students, postdoctoral researchers, and principal investigators as an alpha-test of the web application developed here. Our analysis has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes. Despite this relative conservation, this analysis has also identified many accessions containing new alleles of interest in the *AFB* and *Aux/IAA* families. 
-interesting points from the AFBs
-interesting points IAAs
-TPLs
-ARFs
  -DBD and PB1

Importance of 
-AFB
-IAA degron
-TPL
-ARF DBD and PB1