### 2018-11-15 manuscript/03-methods.Rmd
update viva use outline

# Methods {#methods}

## Data Sources

### Variant Data
Variant data was queried from the 1001 genomes project (http://1001genomes.org) via URL requests to their API service (http://tools.1001genomes.org/api/index.html). These queries provided subsets of the whole-genome VCF. The whole-genome VCF file can be found on the project’s website at http://1001genomes.org/data/GMI-MPI/releases/v3.1/

### Accession Information
Details on each of the 1135 accessions including CS stock numbers, and geographic location where the samples were collected were taken from the 1001 genomes website at http://1001genomes.org/accessions.html, via the download link at the bottom of the page. This data file has been embedded in the R package.

### Gene and transcript information
Information on the genes and transcripts including chromosomal coordinates, start and end location, and transcript length were taken from either the TAIR10 or Araport11 databases. The TAIR10 database, found at http://arabidopsis.org, was accessed via the biomart protocol, using the R package biomaRt. The araport11 database, which can also be found on the arabidopsis.org website (https://www.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FGenes%2FAraport11_genome_release), has been embedded in the R package.

## Nucleotide Diversity Calculation
Nei  defined the nucleotide diversity statistic in his original paper as:
“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
And provided the equation:

\begin{equation}
\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}
(\#eq:nei)
\end{equation}

Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.

a more general form, that treats each sequence in the population as unique can be written as:

\begin{equation}
\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
\qquad
\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
0 \quad \textrm{if} \quad N_{ik} = N_{jk}
\end{array}\right.
(\#eq:nei-gen)
\end{equation}

where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the *i*th sequence of the population.
$L$ is the length of the sequence (the same for all sequences in the population, 
assumes only SNPs exist, no indels). n is the total number of sequences in the population.

from this form we can re-arrange summations to the form below:

\begin{equation}
\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
\qquad
\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}
(\#eq:nei-rearrange)
\end{equation}

here $\pi_k$ can be thought of as the site-wise nucleotide diversity at position k, and is equal to 
the nucleotide diversity of a sequence of length 1 at location k. we can calculate $\pi_k$ for each
site, then average those over the sequence length to calculate $\pi$, the nucleotide diversity 
of the sequence.

The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant. Note, $\pi_k$ is equal to 0 at all locations without variants.
This is also what is displayed in the Diversity Plot tab of the web app.

## Detailed $\pi_k$ calculation simplification.

The formula for $\pi_k$ above requires comparing every sequence to every other sequence at location k,
however, we know there are only a few variant forms at each individual location.

so we can revert back to using Nei's original formula \@ref(eq:nei), modifying it slightly,
replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k.

\begin{equation}
\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = \frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}
\qquad
\pi_{ij(k)} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad i \neq j \\
0 \quad \textrm{if} \quad i = j
\end{array}\right.
(\#eq:nei-simp1)
\end{equation}

Note that in equation \@ref(eq:nei) subscripts i and j are summed over all sequences in the population, however in equation \@ref(eq:nei-simp1) i and j are only summed over unique variants at a particular location k.

we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.

we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}
(\#eq:nei-simp2)
\end{equation}

next we substitute our definition of $n_{!i}$

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})
(\#eq:nei-simp3)
\end{equation}

then distribute and split summation:

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)
(\#eq:nei-simp4)
\end{equation}

finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)
(\#eq:nei-simp5)
\end{equation}

this simplified form for $\pi_k$ is used by the app, because the counts of unique 
variants at a single nucleotide location can easily be summarized in R.

## Software

The r1001genomes package has many software dependencies on other R packages, a few of the key bioinformatics packages used are listed below.

**biomaRt:** The biomaRt package was used for accessing the TAIR10 database on arabidopsis.org

**vcfR:** vcfR was used to read in the VCF files in a flat “tidy” format for easy manipulaiton

**BSgenome:** The BSgenome package BSgenome.Athaliana.TAIR.TAIR9 was used as the source for the complete DNA string of the reference genome (Col-0).

**DECIPHER:** DECIPHER was used to align nucleotide and amino acid sequences of homologous genes

**GenomicFeatures:** The GenomicFeatures package was used for handling sequence annotations.

**Biostrings:** Biostrings provides the underlying framework for the sequence manipulations used for generating and aligning sequences with BSgenome, Decipher, and GenomicFeatures