### 2018-11-15 manuscript/02-results.Rmd
Filled in sections on ViVa app functionality.

# Results and Discussion {#rd}

## An introduction to ViVa

The current version of ViVa is specifically focused on visualizes natural variation in the coding sequences of genes. Future versions will expand upon this to examine noncoding sequences; however, indels are much more common in noncoding sequences, which limits the certainty of this resequencing data.

**Morgan will provide screenshots**

### Data input: Select the desired genes
At the top of the app, below the title but above the tab selection bar are two collapsible panels used for inputting the genes to query and additional annotation for those genes. The first collapsible panel is used to input the desired genes to investigate, either by typing in the TAIR gene IDs, or uploading a file. The second collapsible panel is optionally used to upload annotation files that will be plotted on some of the tabs of the apps second/analysis section.

### SNPstats: Summary of gene information, structure, and diversity
Below the data input section, the rest of the app is divided into several tabs containing the interactive output of the program.

The first tab is the SNP stats tab. This tab provides general information on the gene transcripts being queried, as well as some calculated counts/statistics on the content of variants found in the sample population. The first table of this section is basic information of the transcripts, including TAIR locus and symbol, the chromosomal start and end position of the transcript and the transcript length. This information is collected from the Araport 11 database stored in a local GFF file.

The next two tables provide counts of occurrence of different effect types within the VCF dataset for each transcript. These counts are based on the “Effect” subfield of the “EFF” field in the VCF file. One table provides the total variant counts of each variant type from the reference genome (Col-0) in each accession. The unique allele count table only counts the number of unique variants/alleles, so if multiple accessions have the same variant allele relative the reference, that allele will only be counted once in the unique counts table.

The final table on the SNP Stats tab provides Nei’s diversity statistic for the transcript and the coding region of each gene, as well as the diversity calculated for only synonymous and non-synonymous sites, and the ratio of the non-synonymous to synonymous diversity.

### Diversity plot: Visualize allelic diversity across the coding sequence
The Diversity Plot tab shows Nei’s nucleotide diversity by nucleotide position of the protein coding region of a selected gene. Although the X-axis is marked by codon number from the N-terminus for interpretability, the diversity values are based on single nucleotide sites. The colors of markers on the plot identify the effect classes, based on fields from the VCF. If annotation files are provided, the background of the plot will be color coded by the annotated regions. If points on the plot are selected by clicking and dragging a box over them, the data for the selected points will appear in the grey box below the plot. Below this is a complete data table with the data provided for all points on the plot which can be downloaded as a .csv file.

### Accession mapping
The SNP Mapping tab plots the location of accessions on a world map, and colors the points based on selected variants.

After selecting the genes and filtering on the SNP type and level of nucleotide diversity, a checkbox group becomes available to select variant alleles to display on the map. The variant alleles are labeled with the Transcript_ID and Amino_Acid_Change fields, in the form [Transcript_ID|Amino_Acid_Change].

After selecting the variant alleles and updating the map, the accessions on the map will be grouped and colored by each present unique combination of the selected alleles.
Below the map is a table containing the accession details for all mapped accessions.

### Alignment
The Alignment tab provides DNA and amino acid sequence alignments of selected genes, with loci colored by variants present. The content of this tab is most useful if the selected genes are all family members or have significant sequence homology. If annotation files are uploaded, the background of the sequences will be colored by annotation region. Hovering the mouse over sequence elements with polymorphisms will provide additional details about the SNP present at that loci.

### Gene tree

### SNP browser
The Browser tab provides a way to search and filter the contents of the VCF file by different fields. After selecting the transcripts to include, a number of filters can be applied to the dataset to match text values, or set min and max limits on the values of numeric fields. When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters.

## Visualizing Variation within the auxin signaling pathway

To test the useability and accessibility of this tool we assembled a group of alpha testers comprising postdoctoral, graduate, and undergraduate researchers. This team of researchers analyzed natural variation in the nuclear auxin signaling pathway. This signaling pathway has been well studied, revealing abundant functional knowledge and crystal structures of several protein domains, providing several points of ground truth reference for comparing structure-function variation and natural genetic variation.

Assuming there is selective pressure against mutations which alter a protein's function, nonsynonymous genetic variation in critical functional domains of a protein should be low. Therefore scanning gene coding sequences for regions of low nonsynonymous diversity, may predict functional domains. Within the auxin signaling pathway several domains have been identified as critical to transmitting auxin signals.

Perhaps the most critical domain is the degron of the Aux/IAA transcriptional repressor proteins. Mutations which prevent this domain from being recognized and marked for proteasomal degradation by SCF^TIR1/AFB^ are dominant and often result in obvious phenotypes [@berleth_auxin_2004; @yang_iaa1_2004]. As expected this critical functional region is highly conserved at the amino acid level (\@ref(fig:A-IAA-alignment)). Similarly, the EAR motif which facilitates Aux/IAA interaction with TPL/TPR transcriptional repressors, are also conserved.

```{r Load IAA gene info and VCF}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "IAAs_VCF.rds", package = "r1001genomes"))
```
```{r IAA alignment}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

# make a collapsed alignment for translating positions
collapsed_aln_df <- 
  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
  tidyr::spread(key = seq_name, value = seq_pos)
```
```{r Mask_IAA_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
```
```{bash IAA_phylogeny_construction, eval=FALSE}
~/Desktop/natural-variation-webtool/manuscript

/Applications/MrBayes/mb

execute IAAs.nex
# set IAA33 as the outgroup
outgroup AT5G57420.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```
```{r Load IAA phylogeny and data}
tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")

IAA_data <- read.csv("tables/IAAs_table.csv")
#tip_labels need to be in the first column
IAA_data %<>% dplyr::select(name, everything())
IAA_data$name %<>% as.character()
IAA_data$AGI %<>% as.character()

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```

(ref:A-IAA-alignment) **Canonical IAAs have conserved Degron and EAR motifs.** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious. Red indicates missense variants. Color scale is explained in [Methods](#methods). EAR motif is highlighted by a black box including alignment positions 70-74. Degron domain is highlighted by a gray box including alignment positions 194-201.
```{r A-IAA-alignment, fig.cap="(ref:A-IAA-alignment)"}
A_IAAs <- subset(gene_info, tair_symbol %in% IAA_data[which(IAA_data$Class == "A"), "name"])
A_IAAs <- alignCDS(A_IAAs$transcript_ID)
# make an alignment data frame
A_IAAs <- makeAlnDF(A_IAAs[[2]])
A_IAAs <- left_join(A_IAAs, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
A_IAAs$tair_symbol %<>% as.factor()
A_IAAs$tair_symbol %<>% fct_reorder(
    {as.integer(str_remove(A_IAAs$tair_symbol, 
                           pattern = "IAA"))}, 
    .desc = TRUE)
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
coding_VCF <- getCodingDiv(VCF)
A_IAAs <- addSNPsToAlnDF(aln_df = A_IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
A_IAAs$gap <- A_IAAs$seq_pos == "-"

p <-ggplot(A_IAAs, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = A_IAAs, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("neutral", "deleterious"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p + geom_rect(mapping = aes(xmin = 193, xmax = 202, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA) + geom_rect(mapping = aes(xmin = 69, xmax = 75, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA)
```

The Aux/IAA proteins bind to ARF transcription factors allowing the Aux/IAA's to effect transcriptional repression at specific loci. Thus in the ARF transcription factor proteins we expect the DNA-binding function to be conserved. Indead, the DNA-binding domains of the ARFs had very few, low-diversity missense polymorphisms, as did the PB1 domain which facilitates ARF-Aux/IAA interaction (Figure \@ref(fig:A-ARF-Alignment)).

```{r ARF diversity stats}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "ARF_gene_ids.csv", package = "r1001genomes"))

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "ARF_VCF.rds", package = "r1001genomes"))

VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
VCF <- getCodingDiv(VCF)
```

```{r Mask_ARF_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "ARFs.nex", format = "protein")
```
```{bash ARF_phylogeny_construction, eval=FALSE}
/Applications/MrBayes/mb

execute ARFs.nex
# set ARF17 as the outgroup
outgroup AT1G77850.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```

```{r Load ARF phylogeny and data}
tree <- treeio::read.beast(file = "trees/ARFs.nex.con.tre")

ARF_data <- read.csv(system.file("shiny-app", "gene_family_data",
                      "ARF_gene_ids.csv", package = "r1001genomes"))

#tip_labels need to be in the first column
ARF_data %<>% dplyr::select(name, everything())
ARF_data$tair_locus %<>% as.character()

ARF_data <- ARF_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), ARF_data$tair_locus), ]

tree@phylo$tip.label <- ARF_data$name

```

(ref:A-ARF-Alignment) **Canonical ARFs have conserved DNA binding and PB1 domains.** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious. Red indicates missense variants. Color scale is explained in [Methods](#methods).
```{r A-ARF-Alignment, fig.cap="(ref:A-ARF-Alignment)"}
A_ARFs <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
A_ARFs <- alignCDS(A_ARFs$transcript_ID)
# make an alignment data frame
A_ARFs <- makeAlnDF(A_ARFs[[2]])

A_ARFs <- left_join(A_ARFs, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
#A_ARFs$tair_symbol %<>% factor(levels = ARF_order, 
#                                       ordered = TRUE)
A_ARFs$tair_symbol %<>% fct_rev()

A_ARFs <- addSNPsToAlnDF(A_ARFs, VCF, effect_order = SNPeff_order)
A_ARFs$gap <- A_ARFs$seq_pos == "-"

p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = A_ARFs, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("neutral", "deleterious"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p
#ggsave(p,filename = "A_ARFs_alignment_variants.pdf", width = 12, height = 4)
```

Natural variation also provides a means to study how gene families are evolving. Mapping diversity at nonsynonymous variant sites relative to synonymous sites to gene trees allows comparison of recent functional divergence to the sequence divergence throughout the growth of the gene family. This visualization facilitates hypothesis generation regarding the functional redundancy or novelty of recent gene duplications.

Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005]. For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].

In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved \@ref(fig:IAAtree). In an interspecific comparison between *A. thaliana* and *A. lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017]. Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) \@ref(fig:IAAtree), and also detected high diversity within the same regions of *IAA6* as seen in Winkler *et al*.

(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to infering a phylogeny [@ronquist_mrbayes_2003]. Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$. Nodes are labeled with the poster probability of monophyly.

```{r Load IAA phylogeny and data 2}
tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")

IAA_data <- read.csv("tables/IAAs_table.csv")
#tip_labels need to be in the first column
IAA_data %<>% dplyr::select(name, everything())
IAA_data$name %<>% as.character()
IAA_data$AGI %<>% as.character()

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```
```{r IAAtree, fig.height=5, fig.width=5, fig.cap="(ref:IAAtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
                                        max(IAA_data$PiN_PiS)), 
                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
  theme(legend.position = c(.9,.7)) + 
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
  scale_color_viridis(discrete = TRUE)
p
```

Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(fig:IAAtree) [@remington_contrasting_2004]. Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity. The *Arabidopsis thaliana* EFP browser shows that *IAA10* and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**. In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles. Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.

# Conclusion