### 2018-11-21 manuscript/02-results.Rmd
build after merge

# Results and Discussion {#rd}

## An overview of ViVa

ViVa is meant to visualize natural variation in the coding sequences of genes. For this first iteration, non-coding sequence variation is intentionally excluded from the analysis tools. This reflects challenges both in alignment of non-coding sequences and the increased difficulty in assessing variation in these regions [@alexandre_complex_2018].

```{r}
appFigCaption <- "(ref:app_figure_elements) \n **Key elements of the web application** (a) The first section contains two collapsible panels and is used to input information about the genes to be investigated. (b) The SNP Stats tab provides gene level counts and statistics on SNPs. (c) The Diversity Plot tab plots the nucleotide diversity of SNP sites along the length of the coding region of a selected gene. (d) The SNP Mapping tab plots accessions on a world map colored to indicate what SNPs they have. (e) The  Browser tab allows searching for and filtering variants by text and numeric fields. (f) The Alignments tab aligns DNA and RNA sequences of homologous genes, and colors sequence elements based on SNPs and annotations."
```

```{r, echo = FALSE, fig.cap = appFigCaption, out.width="100%"}
knitr::include_graphics("app_figure/App_figure.png")
```

### Data input: Select the desired genes
At the top of the app, below the title but above the tab selection bar are two collapsible panels used for inputting the genes to query and additional annotation for those genes. The first collapsible panel is used to input the desired genes to investigate, either by typing in the TAIR gene IDs, or uploading a file. The second collapsible panel is optionally used to upload annotation files that will be plotted on some of the tabs of the apps second/analysis section.

### SNPstats: Summary of gene information, structure, and diversity
Below the data input section, the rest of the app is divided into several tabs containing the interactive output of the program.

The first tab is the SNP stats tab. This tab provides general information on the gene transcripts being queried, as well as some calculated counts/statistics on the content of variants found in the sample population. The first table of this section is basic information of the transcripts, including TAIR locus and symbol, the chromosomal start and end position of the transcript and the transcript length. This information is collected from the Araport 11 database stored in a local GFF file.

The next two tables provide counts of occurrence of different effect types within the VCF dataset for each transcript. These counts are based on the “Effect” subfield of the “EFF” field in the VCF file. One table provides the total variant counts of each variant type from the reference genome (Col-0) in each accession. The unique allele count table only counts the number of unique variants/alleles, so if multiple accessions have the same variant allele relative the reference, that allele will only be counted once in the unique counts table.

The final table on the SNP Stats tab provides Nei’s diversity statistic for the transcript and the coding region of each gene, as well as the diversity calculated for only synonymous and non-synonymous sites, and the ratio of the non-synonymous to synonymous diversity.

### Diversity plot: Visualize allelic diversity across the coding sequence
The Diversity Plot tab shows Nei’s nucleotide diversity by nucleotide position of the protein coding region of a selected gene. Although the X-axis is marked by codon number from the N-terminus for interpretability, the diversity values are based on single nucleotide sites. The colors of markers on the plot identify the effect classes, based on fields from the VCF. If annotation files are provided, the background of the plot will be color coded by the annotated regions. If points on the plot are selected by clicking and dragging a box over them, the data for the selected points will appear in the grey box below the plot. Below this is a complete data table with the data provided for all points on the plot which can be downloaded as a .csv file.

### Accession mapping: View distribution of SNPs across the globe
The SNP Mapping tab plots the location of accessions on a world map, and colors the points based on selected variants.

After selecting the genes and filtering on the SNP type and level of nucleotide diversity, a checkbox group becomes available to select variant alleles to display on the map. The variant alleles are labeled with the Transcript_ID and Amino_Acid_Change fields, in the form [Transcript_ID|Amino_Acid_Change].

After selecting the variant alleles and updating the map, the accessions on the map will be grouped and colored by each present unique combination of the selected alleles.
Below the map is a table containing the accession details for all mapped accessions.

### SNP browser: Filter and Search for Variants
The Browser tab provides a way to search and filter the contents of the VCF file by different fields. After selecting the transcripts to include, a number of filters can be applied to the dataset to match text values, or set min and max limits on the values of numeric fields. When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters.

### Alignment: Visualize SNPs on alignments of homologous genes
The Alignment tab provides DNA and amino acid sequence alignments of selected genes, with loci colored by variants present. The content of this tab is most useful if the selected genes are all family members or have significant sequence homology. If annotation files are uploaded, the background of the sequences will be colored by annotation region. Hovering the mouse over sequence elements with polymorphisms will provide additional details about the SNP present at that loci.

### ViVa R package: Programmatic access to ViVa's functionality
All of the functionality of the ViVa web application are implemented through function calls to the ViVa R package. In addition to being able to  generate all figures and tables the web application can, users of the R package gain direct access to the underlying data structures and greater controll over parameters when processing and visualizing that data. The ViVa R package can be found at https://github.com/wrightrc/r1001genomes, or installed by typing the line 
`devtools::install_github("wrightrc/r1001genomes")` into an R console.

## Visualizing Variation within the auxin signaling pathway

To test the useability and accessibility of ViVa, we assembled a group of alpha testers comprising postdoctoral, graduate, and undergraduate researchers at a research university (University of Washington) and at a primarily-undergraduate institutions (Whitman College). Our testers focused their investigation of natural variation on the nuclear auxin signaling pathway. We selected this signaling pathway for multiple reasons including a wealth of functional data and solved structures of several domains or entire proteins. In this way, we were able to assess the predictive ability of the ViVa modules. Summary information about each family examined can be found in the supplement. Below, we describe the results for the Aux/IAA family in more detail.

In most cases, natural selection is expected to minimize the persistence of nonsynonymous mutations in sequences that encode critical functional domains  relative to their persistance in non-critical domains. Therefore, we reasoned that scanning gene coding sequences for regions of relatively low nonsynonymous diversity should highlight functional domains. This general principle can be seen clearly in the analysis of the Aux/IAA family of transcriptional co-repressors/co-receptors. Aux/IAAs have three major domains: domain I facilitates interaction with TPL/TPR transcriptional repressors, domain II facilitates interactions with the TIR1/AFB receptors in the presence of auxin, and the PB1 domain which failitates interactions with the ARF transcription factors. Domain 1 and Domain II can be readily identified (\@ref(fig:A-IAA-alignment)), perhaps because the multiple contributing residues are spread out in linear sequence space. It is worth noting that the most important residues within the PB1 domain also show very little variation.

```{r Load IAA gene info and VCF}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "IAAs_VCF.rds", package = "r1001genomes"))
```
```{r IAA alignment}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

# make a collapsed alignment for translating positions
collapsed_aln_df <- 
  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
  tidyr::spread(key = seq_name, value = seq_pos)
```
```{r Mask_IAA_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
```
```{bash IAA_phylogeny_construction, eval=FALSE}
~/Desktop/natural-variation-webtool/manuscript

/Applications/MrBayes/mb

execute IAAs.nex
# set IAA33 as the outgroup
outgroup AT5G57420.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```
```{r Load IAA phylogeny and data}
tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")

IAA_data <- read.csv("tables/IAAs_table.csv")
#tip_labels need to be in the first column
IAA_data %<>% dplyr::select(name, everything())
IAA_data$name %<>% as.character()
IAA_data$AGI %<>% as.character()

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```

(ref:A-IAA-alignment) **Canonical IAAs have conserved Degron and EAR motifs.** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious. Red indicates missense variants. Color scale is explained in [Methods](#methods). EAR motif is highlighted by a black box including alignment positions 70-74. Degron domain is highlighted by a box including alignment positions 194-201. Critical residues for PB1 domain interactions are highlighted at alignment positions 246, 253, 316, and 320.
```{r A-IAA-alignment, fig.cap="(ref:A-IAA-alignment)"}
A_IAAs <- subset(gene_info, tair_symbol %in% IAA_data[which(IAA_data$Class == "A"), "name"])
A_IAAs <- alignCDS(A_IAAs$transcript_ID)
# make an alignment data frame
A_IAAs <- makeAlnDF(A_IAAs[[2]])
A_IAAs <- left_join(A_IAAs, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
A_IAAs$tair_symbol %<>% as.factor()
A_IAAs$tair_symbol %<>% fct_reorder(
    {as.integer(str_remove(A_IAAs$tair_symbol, 
                           pattern = "IAA"))}, 
    .desc = TRUE)
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
coding_VCF <- getCodingDiv(VCF)
A_IAAs <- addSNPsToAlnDF(aln_df = A_IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
A_IAAs$gap <- A_IAAs$seq_pos == "-"

p <-ggplot(A_IAAs, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = A_IAAs, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("neutral", "deleterious"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p + geom_rect(mapping = aes(xmin = 193, xmax = 202, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA) + geom_rect(mapping = aes(xmin = 69, xmax = 75, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA) + geom_rect(mapping = aes(xmin = 245, xmax = 247, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = .2) +
geom_rect(mapping = aes(xmin = 319, xmax = 321, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = 0.2) +
  geom_rect(mapping = aes(xmin = 252, xmax = 254, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = 0.2) +
  geom_rect(mapping = aes(xmin = 315, xmax = 317, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = 0.2) 
```

Natural variation also provides a means to study how gene families are evolving. To do this, we used ViVa to map the diversity at nonsynonymous variant sites relative to synonymous sites to phylogenetic trees. This visualization makes it straightforward to compare rates of recent functional divergence within the context of rates of sequence divergence across the entire gene family. In this way, a researcher chould hone in on gene clades that are likely to exhibit high rates of functional redundancy or, conversely, gene clades where there is a greater likelihood that novel functions have emerged.

Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005; @winkler_variation_2017]. For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017]. A closer examination of the *IAA19* and *IAA6* pair in *A. thaliana* and *A. lyrata* found that *IAA6* was diversifying far faster than [@winkler_variation_2017]. Consistent with these results, ViVa revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure \@ref(fig:IAAtree)), and also detected high diversity within the same regions of *IAA6* as seen in [@winkler_variation_2017]. This pattern---one *Aux/IAA* sister showing high nonsynonymous diversity while the other sister was more conserved---was observed frequently across the family, suggesting this could be a recurring feature in *Aux/IAA* evolution supporting the large diversity in auxin functions.

(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to infering a phylogeny [@ronquist_mrbayes_2003]. Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$. Nodes are labeled with the poster probability of monophyly.

```{r Load IAA phylogeny and data 2}
tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")

IAA_data <- read.csv("tables/IAAs_table.csv")
#tip_labels need to be in the first column
IAA_data %<>% dplyr::select(name, everything())
IAA_data$name %<>% as.character()
IAA_data$AGI %<>% as.character()

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```
```{r IAAtree, fig.height=5, fig.width=5, fig.cap="(ref:IAAtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
                                        max(IAA_data$PiN_PiS)), 
                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
  theme(legend.position = c(.9,.7)) + 
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
  scale_color_viridis(discrete = TRUE)
p
```

The notable exception to the patter of one conserved and one diersifying sister was observed within the Group B *IAA* genes (*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30*) [@remington_contrasting_2004]. Among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity. In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively). In support of *IAA10* and *IAA11* playing largely redundant roles, the *Arabidopsis thaliana* EFP browser indicates that *IAA10* and *IAA11* have almost identical expression patterns.

# Conclusion