### 2018-12-06 manuscript/02-results.Rmd

# Results and Discussion {#rd}

## An overview of ViVa

ViVa, in this first iteration, is meant to visualize natural variation in the coding sequences of genes. Non-coding sequence variation is intentionally excluded from the analysis tools. This reflects challenges both in alignment of non-coding sequences and the increased difficulty in assessing variation in these regions [@alexandre_complex_2018].

The development version of ViVa can be accessed as a Docker container https://hub.docker.com/r/wrightrc/r1001genomes/ or as an R-package at https://github.com/wrightrc/r1001genomes. ViVa will be hosted at https://www.plantsynbiolab.bse.vt.edu/ViVa/ upon release.

(ref:app-figure-elements) **Key elements of the webtool** (a) The first section contains two collapsible panels, Gene Select and Annotation Files, which are used to input information about the genes to be investigated. (b) The SNP Stats tab provides gene-structure level counts and statistics on SNPs. (c) The Diversity Plot tab plots the nucleotide diversity of SNP sites along the length of the coding region of a selected gene. (d) The SNP Mapping tab plots accessions on a world map colored according to the selected set of SNPs. (e) The SNP Browser tab allows variants and accessions to be filtered by any combination of text and numeric fields. (f) The Alignments tab aligns DNA and amino acid sequences of homologous genes, and colors sequence elements based on SNPs and annotations.
```{r app-figure-elements, fig.cap = "(ref:app-figure-elements)"}
knitr::include_graphics("app_figure/App_figure.png")
```

### Gene Select and Annotation Files  
At the top of the ViVa webtool are two collapsible panels used for entering the genes to query and custom annotations for those genes (Figure \@ref(fig:app-figure-elements)a). The Gene Select panel permits gene input by either typing in the genes' AGI/TAIR locus identifiers, or uploading a .csv file. The Annotation Files panel is optionally used to upload an annotation file containing coordinates of domains, mutations, or any other sequence knowledge that will be plotted on some of the tabs of the webtools analysis section.

Below the data input section, the rest of the webtool is divided into several tabs containing the interactive output of the program.

### SNP Stats: Summary of gene information, structure, and diversity

The SNP Stats tab provides general information on the gene transcripts being queried, as well as calculated counts/statistics on the content of variants found in the sample population (Figure \@ref(fig:app-figure-elements)b). The first table of this section is basic information about the transcripts, including TAIR locus and symbol, the chromosomal start and end position and the transcript length. This information was collected from the Araport11 Official Release (06/2016) annotation dataset [@cheng_araport11_2017].

The next two tables provide counts of SNPs across the gene body for each transcript. The Total Polymorphism Counts table provides the total number of observations of non-reference allele counts of each variant type (the Col-0 accession TAIR9 genome is the reference genome for this dataset). The Unique Allele Count table only counts the number of unique variants and alleles within the population of accessions (*e.g.* if multiple accessions have the same variant allele, that allele will only be counted once).

The Nucleotide Diversity Statistics table provides a nucleotide diversity statistic (${\pi}$) for the transcript and the coding sequence of each gene [@nei_mathematical_1979]. Nucleotide diversity is also calculated for the set of only synonymous (${\pi_S}$) and only non-synonymous sites (${\pi_N}$). The ratio of the presence of non-synonymous to synonymous polymorphism provides a measure of the potential for functional diversity [@firnberg_genetic_2013; @whitehead_optimization_2012]. We present ${\pi_N/\pi_S}$ here as an correlate for functional diversity throughout ViVa [@nelson_snpgenie_2015; @hughes_adaptive_2000]. While imperfect, this metric may be suggestive of functional constraint when ${\pi_N/\pi_S << 1}$ and functional diversity when ${\pi_N/\pi_S >> 1}$ [@hughes_adaptive_1999].

### Diversity plot: Visualize allelic diversity across the coding sequence
The Diversity Plot tab shows the nucleotide diversity of each variant in the coding region of a selected gene (Figure \@ref(fig:app-figure-elements)c). Although the X-axis is marked by codon number from the N-terminus for interpretability, the diversity values are based on single nucleotide sites. The colors of markers on the plot identify the effect of the polymorphism. If annotation files are provided, the background of the plot will be color coded by the annotated regions. If points on the plot are selected by clicking and dragging a box over them, the data for the selected points will appear in the grey box below the plot. Below this is a complete data table containing all points on the plot which can be downloaded as a .csv file. This tab allows users to identify regions of high diversity as well as isolate polymorphisms that may affect gene function and exist in multiple accessions, facilitating phenotypic analysis.

### SNP Mapping: View distributions of SNPs across the globe
The SNP Mapping tab plots the accessions collection locations on a world map, and colors the points based on selected variant alleles (Figure \@ref(fig:app-figure-elements)d). After selecting the genes and filtering on the SNP type and level of nucleotide diversity, a group of checkboxes becomes available to select variant alleles to display on the map. The variant alleles are labeled with the Transcript_ID and Amino_Acid_Change fields, in the form [Transcript_ID|Amino_Acid_Change]. After selecting the variant alleles and updating the map, the accessions on the map will be colored by each present unique combination of the selected alleles. Below the map is a table containing the accession details for all mapped accessions. This tab may help users formulate hypotheses about the relatedness of accessions sharing a common allele and environments in which that allele may be favorable.

### SNP Browser: Filter and search for variants
The SNP Browser tab provides a way to search and filter the variant data by different fields (Figure \@ref(fig:app-figure-elements)e). After selecting the transcripts to include, a number of filters can be applied to the dataset to match text values (e.g. gene name or variant effect), or set minimum and maximum limits on the values of numeric fields (e.g. nucleotide diversity). When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters. This tab can be useful for identifying all accessions with a particular SNP allele, or any non-reference alleles in a particular region of a gene that may not have been easily accessible in another tab.

### Alignments: Visualize SNPs on alignments of homologous genes
The Alignments tab provides DNA and amino acid sequence alignments of selected genes, colored according to the variant allele with the strongest functional effect at each position (Figure \@ref(fig:app-figure-elements)f and \@ref(fig:A-IAA-alignment)). The content of this tab is most useful if the selected genes are all family members or have significant sequence homology. If annotation files are uploaded, the background of the sequences will be colored by annotation region. Hovering the cursor over variants will provide additional details about the alleles present at that locus. This tab facilitates family-wise analysis of functional conservation, allowing users to identify potential functional regions and alleles which may be useful in deciphering this function.

### Gene Tree: Visualize functional diversity and sequence divergence of a gene family
The Gene Tree tab provides a neighbor joining tree (or uploaded tree created by the user) for the selected genes with the tips of the tree mapped with predicted functional diversity as represented by ${\pi_N/\pi_S}$ in the 1001 Genomes dataset (Figure \@ref(fig:IAAtree)). This tab allows users to generate hypotheses regarding functional diversity and redundancy within the context of the predicted evolution of the gene family.

### ViVa R package: Programmatic access to ViVa's functionality
All of the functionalities of the ViVa webtool are implemented through function calls to the ViVa R package. In addition to being able to generate the same sets of figures and tables as in the webtool, users of the R package also gain direct access to the underlying data structures, providing greater control over parameters when processing and visualizing the data. The ViVa R package can be found at https://github.com/wrightrc/r1001genomes and can be installed in your R environment via the `devtools` package: 
`devtools::install_github("wrightrc/r1001genomes")`.

## Visualizing Variation within the auxin signaling pathway

To test the usability and accessibility of ViVa, we assembled a group of alpha testers comprising postdoctoral, graduate, and undergraduate researchers at a research university (University of Washington) and at a primarily-undergraduate institution (Whitman College). Our testers focused their investigation of natural variation on the nuclear auxin signaling pathway. We selected this signaling pathway for multiple reasons including a wealth of functional data and solved structures of several domains or entire proteins. Using this exitsting knowledge, we were able to qualitatively assess the predictive ability of the ViVa modules. Summary information about each nuclear auxin signaling gene family examined can be found in the [Supplemental Data](#si). Below, we describe the results for the *Aux/IAA* family in more detail.

In most cases, natural selection is expected to minimize the persistence of nonsynonymous mutations in sequences that encode critical functional domains relative to their persistence in non-critical domains [@hughes_adaptive_2000]. Therefore, we reasoned that scanning gene coding sequences for regions of relatively low nonsynonymous diversity should highlight functional domains. This general principle can be seen clearly in the analysis of the *Aux/IAA* family of transcriptional co-repressors/co-receptors. Aux/IAAs have three major domains. Domain I contains an EAR motif that facilitates interaction with TOPLESS (TPL) and TOPLESS-related (TPR) transcriptional repressors [@tiwari_aux_2004; @szemenyei_topless_2008]. Domain II, the degron, facilitates interactions with the TIR1/AFB receptors in the presence of auxin [@tan_mechanism_2007]. Domain III (which was originally considered domains III and IV) is a PB1 domain and facilitates interactions with the ARF transcription factors [@ulmasov_aux_1997; @guilfoyle_getting_2012; @nanao_structural_2014; @korasick_molecular_2014]. The EAR motif and degron can be readily identified by the drop in nonsynonymous variation, as visualized by the lack of stronger functional effects in Figure \@ref(fig:A-IAA-alignment). The PB1 domain is not as readily identified, perhaps because the multiple contributing residues are spread out in linear sequence space. It is worth noting that the charged residues that facilitate electrostatic PB1-PB1 interaction show little variation.

```{r Load IAA gene info and VCF}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "IAAs_VCF.rds", package = "r1001genomes"))
```
```{r IAA alignment}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

# make a collapsed alignment for translating positions
collapsed_aln_df <- 
  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
  tidyr::spread(key = seq_name, value = seq_pos)
```
```{r Mask_IAA_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
```
```{bash IAA_phylogeny_construction, eval=FALSE}
~/Desktop/natural-variation-webtool/manuscript

/Applications/MrBayes/mb

execute IAAs.nex
# set IAA33 as the outgroup
outgroup AT5G57420.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```
```{r Load IAA phylogeny and data}
tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")

IAA_data <- read.csv("tables/IAAs_table.csv")
#tip_labels need to be in the first column
IAA_data %<>% dplyr::select(name, everything())
IAA_data$name %<>% as.character()
IAA_data$AGI %<>% as.character()

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```

(ref:A-IAA-alignment) **Canonical Aux/IAAs have conserved Degron and EAR motifs.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods).

```{r A-IAA-alignment, fig.cap="(ref:A-IAA-alignment)"}
A_IAAs <- subset(gene_info, tair_symbol %in% IAA_data[which(IAA_data$Class == "A"), "name"])
A_IAAs <- alignCDS(A_IAAs$transcript_ID)
# make an alignment data frame
A_IAAs <- makeAlnDF(A_IAAs[[2]])
A_IAAs <- left_join(A_IAAs, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
A_IAAs$tair_symbol %<>% as.factor()
A_IAAs$tair_symbol %<>% fct_reorder(
    {as.integer(str_remove(A_IAAs$tair_symbol, 
                           pattern = "IAA"))}, 
    .desc = TRUE)
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
coding_VCF <- getCodingDiv(VCF)
A_IAAs <- addSNPsToAlnDF(aln_df = A_IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
A_IAAs$gap <- A_IAAs$seq_pos == "-"

p <-ggplot(A_IAAs, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = A_IAAs, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = expand_scale(add = c(0,2.5))) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", 
       fill = "variant functional effect", title = "") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p + annotate("rect", xmin = 69, xmax = 75, ymin = 0.4, 
             ymax = length(levels((A_IAAs$tair_symbol))) + 0.6, 
             color= "black", fill = NA) + 
  annotate("text", x = 72, y = length(levels((A_IAAs$tair_symbol))) + 1.2, 
                          label = "EAR domain", family="Arial") + 
  annotate("rect", xmin = 193, xmax = 202, ymin = 0.4, 
           ymax = length(levels(A_IAAs$tair_symbol)) + 0.6, color= "black", 
           fill = NA) + 
  annotate("text", x = (202-193)/2+193, 
           y = length(levels(A_IAAs$tair_symbol)) + 1.2, label = "degron", 
           family="Arial") + 
  annotate("rect", xmin = 245, xmax = 247, ymin = 0.4, 
           ymax = length(levels(A_IAAs$tair_symbol))+0.6, color = "black", 
           fill = NA) +
  annotate("rect", xmin = 252, xmax = 254, ymin = 0.4, 
           ymax = length(levels(A_IAAs$tair_symbol)) + 0.6, color = "black", 
           fill = NA) +
  annotate("rect", xmin = 315, xmax = 317, ymin = 0.4, 
           ymax = length(levels(A_IAAs$tair_symbol)) + 0.6, color = "black", 
           fill = NA) +
  annotate("rect", xmin = 319, xmax = 321, ymin = 0.4, 
           ymax = length(levels(A_IAAs$tair_symbol)) + 0.6, color = "black", 
           fill = NA) + 
  annotate("text", x = (321-245)/2+245, y= length(levels(A_IAAs$tair_symbol)) + 1.2, label = "PB1 domain", family="Arial")

#ggsave("A-IAAs_alignment.png", width = 6, height = 4)
```

Natural variation also provides a means to study how gene families are evolving. To do this, we used ViVa to map onto the Aux/IAA phylogenetic tree the diversity at nonsynonymous variant sites relative to synonymous sites (Figure \@ref(fig:IAAtree)). This visualization enables straightforward comparison of rates of recent functional divergence within the context of rates of sequence divergence across the entire gene family. By comparing nonsynonymous diversity gene clades can be identified that are likely to exhibit high rates of functional conservation and possible redundancy or conversely, where there is a possibility for recent emergent novel function or pseudogenization.

Previous research has found evidence of both broad genetic redundancy among the *Aux/IAAs* and also specificity within closely related pairs or groups of Aux/IAA proteins [@overvoorde_functional_2005; @winkler_variation_2017]. For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/IAA19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017]. A closer examination of the *IAA19* and *IAA6* pair within *Brassicaceae* found evidence for positive selection and possible subfunctionalization of *IAA6* relative to *IAA19* [@winkler_variation_2017]. Consistent with these results, ViVa revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure \@ref(fig:IAAtree)), and also detected high nonsynonymous diversity within the same regions of *IAA6* as seen by Winkler et al. (Figure \@ref(fig:IAA6-div-plot)). This pattern---one sister showing high nonsynonymous diversity while the other sister was more conserved---was observed frequently across the *Aux/IAA* as well as the *AFB* and *ARF* families (Figure \@ref(fig:AFBtree) and \@ref(fig:ARFtree)), suggesting this could be a recurring feature in the evolution of these families supporting the large diversity in auxin functions.

(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to inferring a phylogeny [@ronquist_mrbayes_2003]. Tips of the tree are mapped with circles of color and diameter proportional to ${\pi_N/\pi_S}$. ${\pi_N/\pi_S}$ provides a prediction of functional diversity. Nodes are labeled with the poster probability of monophyly. There are two distinct clades of Aux/IAAs represented by the majority of the A and B classes. C class Aux/IAAs are missing one or more of the canonical Aux/IAA domains.

```{r Load IAA phylogeny and data 2}
tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")

IAA_data <- read.csv("tables/IAAs_table.csv")
#tip_labels need to be in the first column
IAA_data %<>% dplyr::select(name, everything())
IAA_data$name %<>% as.character()
IAA_data$AGI %<>% as.character()

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```
```{r IAAtree, fig.height=5.5, fig.width=5, fig.cap="(ref:IAAtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
                                        max(IAA_data$PiN_PiS)), 
                                      to = c(0,1)), guide = "colourbar", breaks = c(min(IAA_data$PiN_PiS), 1,
                                        max(IAA_data$PiN_PiS)), labels = c(min(IAA_data$PiN_PiS), 1,
                                        max(IAA_data$PiN_PiS))) + scale_size(guide = "none") + 
  theme(legend.position = c(.8,.7)) + 
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
  scale_color_viridis(end = .9, discrete = TRUE, option = "A") +  
  labs(fill = expression(pi [N] / pi [S]), color = "IAA class")
p

#ggsave("IAA-tree.png", width = 6, height = 5.5)
```

```{r IAA-tree-order}
IAA_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)
```

The *Aux/IAA* phylogeny clusters into two distinct clades represented by the A and B classes [@remington_contrasting_2004]. The C class Aux/IAAs are missing one or more of the canonical Aux/IAA domains. We found notable exceptions to pattern of diversification and conservation between sister pairs within the Class B *Aux/IAA* genes. The *IAA10*/*IAA11*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* pairs showed similar levels of nonsynonymous diversity. For example, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively). In support of this strong conservation of both *IAA10* and *IAA11*, the *Arabidopsis thaliana* ePlant browser indicates that *IAA10* and *IAA11* have almost identical expression patterns [@waese_eplant_2017]. Together this evidence suggests a strong dosage requirement for these genes or that they have taken on novel functions since their emergence.